---
date     : "2017-04-26T11:56:32+08:00"
draft    :  false
title    : "Build a R Package with Rcpp in 30 minutes"
subtitle : "Example : Poisson LASSO Regression" 
fontsize :  18pt
---
\
\

<p style=' font-family:"Times New Rome"; font-size:14pt '>
<code> Rcpp </code> is package with which you could write c++ code within R. 
</p>

<p style=' font-family:"Times New Rome"; font-size:14pt '>
You could simply write a inline C++ code in R like this:
</p>

\
```{r, echo=TRUE, eval=FALSE}
library(Rcpp)

# Define a Rcpp function
add_vec <- Rcpp::cppFunction(
  
  'Rcpp::NumericVector add_vec( Rcpp::NumericVector a, Rcpp::NumericVector b){
  
  return a+b;
  
  }'
  
)

# See the result
t(add_vec(c(1,2,3), c(2,3,4)))

```
\

<p style=' font-family:"Times New Rome"; font-size:14pt '>
Or you could save your <code> .cpp </code> file and call the function from R. With this
powerful tool, you gain the productivity of R and the speed of C++.
</p>

<p style=' font-family:"Times New Rome"; font-size:14pt '>
Let's try to build a small package with <code> Rcpp </code>. Take the Poisson LASSO regression as an example. I'm gonna to install this model with ADMM algorithm and through <code> RcppArmadillo </code> package.
</p>

<p style=' font-family:"Times New Rome"; font-size:14pt '>
So let's get start:
</p>

### __Prerequisite__

<dl style=' font-family:"Times New Rome"; font-size:14pt '>
[//]:# (This is just a comment test)

+ <code> R </code> 
+ <code> Rcpp </code>
+ <code> RcppArmadillo </code>
</dl>

### __Build the Skeleton of the Package__

<code> Rcpp </code> provides function that could help you build a package very quickly. Execute next line will build a package skeleton named "PoisLASSO" in you working directory:

```{r, echo=TRUE, eval=FALSE}

Rcpp.package.skeleton('PoisLASSO') # Your package name should go there

```

<p style=' font-family:"Times New Rome"; font-size:14pt '>
In the folder PoisLASSO, there is a folder called src where you should put all your C++ files there. Let's build a new C++ file called "*PoisLASSO.cpp*". And we'll put our function in this file.
</p>

### __Poisson LASSO Regression and ADMM algorithm__

<p style=' font-family:"Times New Rome"; font-size:14pt '>
I'll skip the detail of this algorithm. But the main idea is to iteratively optimize the objective function and its conjugate problem. The objective function is solved with Lagrange multipler method with augmented variable. And the algorithm is presented as below:
</p>

\
```{r, echo=TRUE, engine='Rcpp', eval=FALSE}
  
  #include <iostream>
  #include <RcppArmadillo.h>
  //[[Rcpp::depends(RcppArmadillo)]]
  using namespace Rcpp;
  using namespace arma;
  using namespace std;
  
  arma::vec ADMM_Lasso( arma::mat X, arma::colvec Y, double lambda, double eta){

  int nVar  = X.n_cols;
  int marker = 1;
  double tol   = 0.01;
  double tol_1 = 0.1;
  double tol_2 = 0.1;
  int terminate = 0;
  int converge  = 0;

  // Initialization
  arma::vec b_gd    = rnorm(X.n_cols, 0, 1) ;
  arma::vec z_admm  = rnorm(X.n_cols, 0, 1) ;
  arma::vec z_old = z_admm;
  arma::vec mu_admm = rnorm(X.n_cols, 0, 1) ;
  // Using ADAM GD method
  NumericVector m_temp;
  NumericVector v_temp;
  arma::vec m;
  arma::vec v;
  arma::vec steps;
  
  while(!terminate){
    m_temp = rep(0,X.n_cols);
    m = as<arma::vec>(m_temp); 
    v_temp = rep(0,X.n_cols);
    v = as<arma::vec>(v_temp);
    converge  = 0;
    int marker_inner = 0;
    
    // Optimize Beta
    while(!converge){
      arma::vec G_1(X.n_cols);
      arma::vec G_2(X.n_cols);
      arma::vec G_3(X.n_cols);
      arma::vec G(X.n_cols);
      for(int i=0; i<nVar; i++){
        G_1(i) = sum(X.col(i) % Y);
        G_2(i) = sum(exp(X * b_gd) % X.col(i));
      }
      G_3 = (b_gd - z_admm) + mu_admm;
      G = G_1 - G_2 + G_3;
      m = eta * 0.1*m + eta * 0.9*G;
      v = 0.1*v + 0.9*G % G;
      steps = m % pow(v, -0.5);
      b_gd  = b_gd  +  steps;
      Rcout << "b_gd : " << trans(b_gd) << std::endl;
      marker_inner = marker_inner + 1;
      
      if( max(steps) < tol || marker_inner > 20 ){
        converge = 1;
        break;
      }
    }

    // Soft Threshhold Z
    for(int i=0; i<nVar; i++){
      if(b_gd(i)+mu_admm(i) > lambda){
        z_admm(i) = b_gd(i) + mu_admm(i) - lambda;
      }
      else if(b_gd(i)+mu_admm(i) < -lambda){
        z_admm(i) = b_gd(i) + mu_admm(i) + lambda;
      }
      else{
        z_admm(i) = 0;
      }
    }
    
    // For monitor reason
    Rcout << "z_admm : "  << trans(z_admm) << endl;
    // Updating augmented mu
    mu_admm = b_gd - z_admm + mu_admm;
    // For monitor reason
    Rcout << "mu_admm : " << trans(mu_admm) << endl;
    // Check Convergence
    int cond_1 = (max(abs(b_gd - z_admm))  < tol_1);
    int cond_2 = (max(abs(z_admm - z_old)) < tol_2);
    if((cond_1==1) && (cond_2==1)){
      terminate = 1;
      break;
    }

    z_old = z_admm;
    marker = marker + 1;
    Rcout << "Iteration : " << marker << std::endl;
  }

  Rcpp::Rcout << trans(b_gd)    << std::endl ;
  Rcpp::Rcout << trans(z_admm)  << std::endl ;
  Rcpp::Rcout << trans(mu_admm) << std::endl ;
  return(b_gd);
}

```


### __Build the Package and Upload to Github__

<p style=' font-family:"Times New Rome"; font-size:14pt '>
Go to the upper directory of your current directory, which contain the PoissLASSO folder. And open your command line (you could do it with <code> devtools </code> as well), and write:
</p>

\
```{bash, eval=FALSE, echo=TRUE}
R CMD build PoissLASSO
R CMD check PoissLASSO
```
\

<p style=' font-family:"Times New Rome"; font-size:14pt '>
If everything is OK, you could now install your package locally <code> devtools::install('PoissLASSO') </code>.
</p>

<p style=' font-family:"Times New Rome"; font-size:14pt '>
And upload your package to Github with:
</p>

\
```{bash, eval=FALSE, echo=TRUE}
git init
git add .
git remote add -v "Your Repo URL go here"
git commit -m "First Commit"
git push -u origin master
```
\

<p style=' font-family:"Times New Rome"; font-size:14pt '>
And after finishing this, you should be able to download your own package with 
</p>

\
```{r, echo=TRUE, eval=FALSE}
devtools::install_github('Your Git Account Name/Your repo name')
library('PoissLASSO')
```
\




